### 2022-03-05 tex/manuscript.tex
Edit introdu nullcline description

\documentclass[utf8, draft]{frontiersFPHY} % for Physics and Applied Maths
% \documentclass[utf8]{frontiersFPHY} % for Physics and Applied Maths
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{soul}
\usepackage{physics}
\usepackage{cases}
\usepackage{bm}
% This should make nicer typesetting and lead to feaver underfull/overfull box errors
\usepackage{microtype}

\usepackage{url,hyperref,microtype,subcaption}
% \usepackage[onehalfspacing]{setspace}
\usepackage{cleveref}
% \usepackage[font={bold,it}]{caption}

\graphicspath{{../figures/}}

% Define better ref abbreviations
\usepackage{lineno}

% Define terms for "free"/"quiet"
\newcommand{\free}{free}
\newcommand{\quiet}{suppressed}

% Define some convenience commands
\renewcommand{\d}{\mathrm{d}}
\renewcommand{\k}{\mathrm{K}}
\newcommand{\ca}{\mathrm{Ca}}
\newcommand{\na}{\mathrm{Na}}
\newcommand{\leak}{\mathrm{L}}
\newcommand{\dstar}{d^\star}
\newcommand{\gstar}{g^\star}
\newcommand{\gbar}{\bar g}
\newcommand{\delt}{\Delta t}
\newcommand{\taus}{\tau_s}
\newcommand{\dn}{\delta_n}
\newcommand{\taud}{\tau_d}

\linenumbers

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Olenik {et~al.}} %use et al only if is more than 1 author
\def\Authors{Mark Olenik\,\(^{1,*}\), and Conor Houghton\,\(^{2}\)}
\def\Address{\(^{1}\)School of Biological Sciences, Faculty of Life Sciences, University of Bristol, Bristol, United Kingdom \\
\(^{2}\)School of Computer Science, Electrical and Electronic Engineering, and Engineering Mathematics, Faculty of Engineering, University of Bristol, Bristol, United Kingdom}
\def\corrAuthor{Mark Olenik}

\def\corrEmail{m.olenik@bristol.ac.uk}

\begin{document}

\onecolumn
\firstpage{1}

\title[Scalar Poincaré map for bursting]{A scalar Poincaré map for anti-phase bursting in coupled inhibitory neurons with synaptic depression.}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.

\maketitle

\begin{abstract}
  \section{}
  Short-term synaptic plasticity is found in many areas of the central nervous system.
  In the inhibitory half-centre central pattern generators involved in locomotion, synaptic depression is believed to act as a burst termination mechanism, allowing networks to generate anti-phase bursting patterns of varying periods.
  To better understand burst generation in these central patter generators, we study a minimal network of two neurons coupled through depressing synapses.
  Depending on the strength of the synaptic conductance between the two neurons, this network can produce symmetric \(n-n\) anti-phase bursts, where neurons fire \(n\) spikes in alternation, with the period of such solutions increasing with the strength of the synaptic conductance.
  Relying on the timescale disparity in the model, we reduce the eight-dimensional network equations to a fully-explicit scalar Poincaré burst map.
  This map tracks the state of synaptic depression from one burst to the next and captures the complex bursting dynamics of the network.
  Fixed points of this map are associated with stable burst solutions of the full network model, and are created through fold bifurcations of maps.
  We derive conditions that describe period-increment bifurcations between stable \(n-n\) and \((n+1)-(n+1)\) bursts, producing a full bifurcation diagram of the burst cycle period.
  Predictions of the Poincaré map fit excellently with numerical simulations of the full network model and allow the study of parameter sensitivity for rhythm generation.

  \tiny
  \keyFont{ \section{Keywords:} Synaptic depression, Poincaré map, Dynamical system, Neuronal bursting, Central pattern generator}

\end{abstract}

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/discussion}

\section*{Conflict of Interest Statement}
The authors declare that the research was conducted in the absence of any commercial or
financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
MO and CH contributed to conception and design of the study.
MO performed numerical computation and analysis.
All authors contributed to manuscript writing and revision.

\section*{Funding}
This work was supported by the Wellcome Trust Doctoral Training Programme in Neural
Dynamics, Grant no. 099699/Z/12/Z.

\section*{Acknowledgements}
The first author thanks the Wellcome Trust for financial support of his Ph.D. study.
We thank Alan Champneys for sharing his insights on non-continuous maps during the course of this research.
We are also grateful to Alan Roberts and Stephen R. Soffe for their comments on earlier versions of the manuscript.

\section*{Data Availability Statement}
The datasets generated for this study are available on request to the corresponding author.

\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{bibliography.bib}

\section*{Figure captions}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/nullclines-uncoupled.pdf}
  \caption{Periodic solution of relaxation-oscillator model neuron. \(\bm{\mathrm{(A)}}\)
    Projection of limit cycle onto (\(v,w\))-phase plane with \(v\)-nullcline (blue,
    \(v_\infty\)) and \(w\)-nullcline (orange, \(w_\infty\)). Unstable fixed point \(p_{f}\)
    is indicated by an orange dot. \(\bm{\mathrm{(B)}}\) Corresponding voltage trace
    \(v(t)\) of an action potential.~\label{fig:nullclines}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/depression-traces.pdf}
  \caption{Solution profiles of a \(4-4\) burst. \(\bm{\mathrm{(A)}}\) Membrane potentials
    of cell 1 (\(v_{1}\), blue), and cell 2 (\(v_{2}\), orange). \(\bm{\mathrm{(B)}}\)
    Synaptic variables \(d_{1}\) (blue) and \(s_{1}\) (grey) of cell
    1.~\label{fig:depression-traces}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/burst-sols.pdf}
  \caption{Voltage traces of both cells of numerically stable solutions for increasing
    values of the coupling strength \(\gbar\) (increasing top to
    bottom).~\label{fig:burst-sols}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/bif-diagram.pdf}
  \caption{Numerically computed bifurcation diagrams of stable \(n-n\) solutions for
    increasing coupling strength \(\gbar\). \(\bm{\mathrm{(A)}}\) Period of stable
    solutions. Dashed lines show the interval between the \(5-5\) and the suppressed
    solutions, where higher period \(n-n\) solutions occur on increasingly smaller intervals
    of \(\gbar\). \(\bm{\mathrm{(B)}}\) \(ISI\)s corresponding to \(n-n\) solutions in
    \(\mathrm{(A)}\). Long \(ISI\)s are associated with the quiet phase of a burst, short
    \(ISI\)s with the free phase. During the free phase, \(ISI\)s are of approximately
    constant duration \(T\).\label{fig:bif-diagram}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/nullclines-coupled.pdf}
  \caption{Nullclines \(v_{\infty}\) (blue) and \(w_{\infty}\) (orange) in the
  \((v,w\))-phase plane for different values of the total synaptic conductance \(\gbar
  s\). For small \(\gbar s < g^{\star}\), fixed point \(p_{f}\) is unstable (orange
  point). Larger values \(\gbar s\) move \(v_{\infty}\) down in the \((v,w)\)-plane until
  \(p_{f}\) changes stability (half orange,  half blue point) at some critical total
  conductance value \(\gbar s = g^{\star}\), and becomes stable (blue point) for
  \(\gbar>g^{\star}\).~\label{fig:nullclines-coupled}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\textwidth]{fig/release-delay.pdf}
  \caption{Numerically computed values of the release delay for varying \(\gbar\). Each of
    the three branches (A, B, C) also shows the timecourse of the total synaptic
    conductance \(\gbar s\) of a sample stable solution of both cells (blue and orange),
    as well as the release conductance \(g^{\star}\) (dashed green line).
    \(\bm{\mathrm{(A)}}\) Branch corresponding to the \(1-1\) solution. Here the quiet
    cell only spikes after a significant release delay. \(\bm{\mathrm{(B)}}\) Branch with
    a long release delay associated with a subset of \(2-2\) solutions. Here the release
    condition is briefly satisfied after the first spike of cell 1. This does not cause
    firing of cell 2, which only occurs after the second spike of cell 1.
    \(\bm{\mathrm{(C)}}\) Branch with \(n-n\) solutions  where release delay is
    approximately zero and the release condition is sufficient for firing of cell 2.
    ~\label{fig:release-delay}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/free-quiet.pdf}
  \caption{Schematic diagram of the free and quiet phases for a \(3-3\) solution.
    \(\bm{\mathrm{(A)}}\) Membrane potentials of cell 1 (\(v_{1}\)) and cell 2 (\(v_{2}\)).
    The grey patches depict inter-burst-intervals \(\delt\). \(\bm{\mathrm{(B)}}\) Total
    synaptic conductance of cell 1 (\(\gbar s_1\)) as it crosses the release conductance
    \(g^{\star}\). \(\bm{\mathrm{(C)}}\) Solution \(d_1(t)\) of depression variable of cell
    1, during free (blue) and quiet phases (orange).~\label{fig:free-quiet1}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/delta-t.pdf}
  \caption{Numerically computed bifurcation diagram of \(\delt\) for varying \(\gbar\).
    Each continuous branch is associated with a stable \(n-n\) burst solution. Increasing
    \(\gbar\) increases \(\Delta t\) until the solutions bifurcate at \(\Delta t\approx
    T\).~\label{fig:delta-t}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/FQ-map.pdf}
  \caption{Maps \(F_n\) \(\bm{\mathrm{(A)}}\) and \(Q_n\) \(\bm{\mathrm{(B)}}\) for
  \(\gbar=0.6\) \(\si{mS/cm^{2}}\) and \(n=1,2,3,4\). Curves \(F_n\) intersect at
  \(d_{s}\) which is indicated by a dashed vertical line.~\label{fig:FQ-map}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/Pn-map.pdf}
  \caption{Map \(\Pi_{n}:d^{\star}\). \(\bm{\mathrm{(A)}}\) \(\Pi_{n}\) for \(n=1,2,3,4\)
  at \(\gbar=0.6\) \(\si{mS/cm^{2}}\). \(\bm{\mathrm{(B)}}\) \(\Pi_{2}\) with \(n=2\) for
  varying \(\gbar \approx 0.01, 0.034, 0.3\) \(\si{mS/cm^{2}}\). The identity function is
  illustrated by a diagonal line.~\label{fig:Pn-map}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{fig/folds.pdf}
  \caption{\(\bm{\mathrm{(A)}}\) Fold bifurcation diagrams of stable (continuous curves)
    and unstable (dotted curves) fixed points of \(\Pi_{n}\) for varying \(n\).
    \(\bm{\mathrm{(B)}}\) Cycle periods computed from stable fixed points (blue), and the
    corresponding solution period from numerical integration of the system of ODEs
    (orange).~\label{fig:folds}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\textwidth]{fig/final-bif.pdf}
  \caption{Bifurcation diagrams of stable \(n-n\) solutions computed analytically from
    fixed points of \(\Pi_n\) and plotted on the respective intervals of \(\gbar\in
    \big[\gbar_+(n),\gbar_-(n)\big]\) (blue), and computed from numerical integrations of
    the ODEs (orange).~\label{fig:final-bif}}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: