### 2022-03-07 tex/response.tex
Refactor a bunch of plots

\documentclass[utf8]{frontiers_suppmat} % for all articles
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\usepackage{cleveref}
\usepackage{physics}
\usepackage{siunitx}
\usepackage{xr}
\graphicspath{{../figures/}}
\externaldocument{draft}

\renewcommand{\d}{\mathrm{d}}
\renewcommand{\k}{\mathrm{K}}
\newcommand{\ca}{\mathrm{Ca}}
\newcommand{\na}{\mathrm{Na}}
\newcommand{\leak}{\mathrm{L}}
\newcommand{\dstar}{d^\star}
\newcommand{\gstar}{g^\star}
\newcommand{\gbar}{\bar g}
\newcommand{\delt}{\Delta t}
\newcommand{\taus}{\tau_s}
\newcommand{\dn}{\delta_n}
\newcommand{\taud}{\tau_d}

% Leave a blank line between paragraphs instead of using \\

\begin{document}
\onecolumn
\firstpage{1}

\title[Response]{{\helveticaitalic{Response}}}
\maketitle

% TODO: Add line numbers everywhere.
\section{Minor changes}
\subsection{Changes to the nullclines plot}
In order to reduce the number of figures we have merged figures 1 and 5, and adjusted added a paragraph describing the non-zero constant synaptic conductance case at the end of the methods part.

\section{Release delay}
Due to the changes to the single cell model parameters, i.e. reducing the applied current from $I=6 \si{\mu A/cm^{2}}$ to $I=3.8\si{\mu A/cm^{2}}$, the $v$-nullcline moved down along the $w$-nullcline.
% I've checked that old HB value manually, should be correct.
This made the system become more excitable, that is the subcritical Andronov-Hopf bifurcation point where the fixed point changes stability moved from $\bar g=0.038$ to $\bar g=0.0038$.
As a consequence, the release delay seen previously is now negligible for all $n:n$ solutions, see figure \cref{fig:release-delay}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\textwidth]{release-delay.pdf}
  \caption{Numerically computed graph of the release delay as a function of $\bar g$~\label{fig:release-delay}}
\end{figure}

\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{bibliography.bib}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: