### 2022-03-30 tex/manuscript.tex

\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Maths
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{soul}
\usepackage{physics}
\usepackage{cases}
\usepackage{bm}
% This should make nicer typesetting and lead to feaver underfull/overfull box errors
\usepackage{microtype}

\usepackage{url,hyperref,microtype,subcaption}
% \usepackage[onehalfspacing]{setspace}
\usepackage{cleveref}
% \usepackage[font={bold,it}]{caption}

\graphicspath{{../figures}}

% Define better ref abbreviations
\usepackage{lineno}

\usepackage{xr}
\externaldocument{supplementary}

% Define terms for "free"/"quiet"
\newcommand{\free}{active}
\newcommand{\suppressed}{silent}

% Define some convenience commands
\renewcommand{\d}{\mathrm{d}}
\renewcommand{\k}{\mathrm{K}}
\newcommand{\ca}{\mathrm{Ca}}
\newcommand{\na}{\mathrm{Na}}
\newcommand{\leak}{\mathrm{L}}
\newcommand{\dstar}{d^\star}
\newcommand{\gstar}{g^\star}
\newcommand{\gbar}{\bar g}
\newcommand{\delt}{\Delta t}
\newcommand{\taus}{\tau_s}
\newcommand{\dn}{\delta_n}
\newcommand{\taud}{\tau_d}

\linenumbers

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Olenik {et~al.}} %use et al only if is more than 1 author
\def\Authors{Mark Olenik\,$^{1,*}$, and Conor Houghton\,$^{2}$}
\def\Address{$^{1}$School of Biological Sciences, Faculty of Life Sciences, University of Bristol, Bristol, United Kingdom \\
$^{2}$School of Computer Science, Electrical and Electronic Engineering, and Engineering Mathematics, Faculty of Engineering, University of Bristol, Bristol, United Kingdom}
\def\corrAuthor{Mark Olenik}

\def\corrEmail{m.olenik@bristol.ac.uk}

\begin{document}

\onecolumn
\firstpage{1}

\title[Scalar Poincaré map for bursting]{A scalar Poincaré map for anti-phase bursting in coupled inhibitory neurons with synaptic depression.}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.

\maketitle

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/discussion}

\section*{Conflict of Interest Statement}
The authors declare that the research was conducted in the absence of any commercial or
financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
MO and CH contributed to conception and design of the study.
MO performed numerical computation and analysis.
All authors contributed to manuscript writing and revision.

\section*{Funding}
This work was supported by the Wellcome Trust Doctoral Training Programme in Neural
Dynamics, Grant no. 099699/Z/12/Z.

\section*{Acknowledgements}
The first author thanks the Wellcome Trust for financial support of his Ph.D. study.
We thank Alan Champneys for sharing his insights on non-continuous maps during the course of this research.
We are also grateful to Alan Roberts and Stephen R. Soffe for their comments on earlier versions of the manuscript.

\section*{Data Availability Statement}
The data sets generated for this study are available on request to the corresponding author.

\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{bibliography.bib}

\section*{Figure captions}

\begin{figure}[h!]
  \centering
  \includegraphics{nullclines}
  \caption{Periodic solution of relaxation-oscillator model neuron. $\bm{\mathrm{(A)}}$
    Projection of limit cycle onto ($v,w$)-phase plane with $v$-nullcline (blue,
    $v_\infty$) and $w$-nullcline (orange, $w_\infty$). Unstable fixed point $p_{f}$
    is indicated by an orange dot. $\bm{\mathrm{(B)}}$ Corresponding voltage trace
    $v(t)$ of an action potential.~\label{fig:nullclines}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{burst-sols}
  \caption{Voltage traces of both cells of numerically stable solutions for increasing
    values of the coupling strength $\gbar$ (increasing top to
    bottom).~\label{fig:burst-sols}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{bif-diagram}
  \caption{Numerically computed bifurcation diagram of the cycle period of stable $n:n$ solutions for increasing coupling strength $\gbar$. Dashed lines show the interval between the $5-5$ and the suppressed solution, where higher period $n:n$ solutions occur on increasingly smaller intervals of $\gbar$.\label{fig:bif-diagram}}
\end{figure}

% \begin{figure}[h!]
%   \centering
%   \includegraphics{nullclines-coupled}
%   \caption{Nullclines $v_{\infty}$ (blue) and $w_{\infty}$ (orange) in the
%   $(v,w$)-phase plane for different values of the total synaptic conductance $\gbar
%   s$. For small $\gbar s < g^{\star}$, fixed point $p_{f}$ is unstable (orange
%   point). Larger values $\gbar s$ move $v_{\infty}$ down in the $(v,w)$-plane until
%   $p_{f}$ changes stability (half orange,  half blue point) at some critical total
%   conductance value $\gbar s = g^{\star}$, and becomes stable (blue point) for
%   $\gbar>g^{\star}$.~\label{fig:nullclines-coupled}}
% \end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{gstar-diag}
  \caption{Numerically computed values of the release conductance ($\tilde g^{\star}$) for various $n:n$ solutions. The dashed line indicates the analytical approximation of the release conductance by $g^{\star}$.\label{fig:gstar-diag}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{free-quiet}
  \caption{Schematic diagram of the free and quiet phases for a $3-3$ solution.
    $\bm{\mathrm{(A)}}$ Membrane potentials of cell 1 ($v_{1}$) and cell 2 ($v_{2}$).
    The grey patches depict inter-burst-intervals $\delt$. $\bm{\mathrm{(B)}}$ Total
    synaptic conductance of cell 1 ($\gbar s_1$) as it crosses the release conductance
    $g^{\star}$. $\bm{\mathrm{(C)}}$ Solution $d_1(t)$ of depression variable of cell
    1, during free (blue) and quiet phases (orange).~\label{fig:free-quiet}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{delta-t}
  \caption{Numerically computed bifurcation diagram of $\delt$ for varying $\gbar$.
    Each continuous branch is associated with a stable $n:n$ burst solution. Increasing
    $\gbar$ increases $\Delta t$ until the solutions bifurcate at $\Delta t\approx
    T$.~\label{fig:delta-t}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{FQ-maps}
  \caption{Maps $F_n$ $\bm{\mathrm{(A)}}$ and $Q_n$ $\bm{\mathrm{(B)}}$ for
  $\gbar=0.5$ $\si{mS/cm^{2}}$ and $n=1,2,3,4$. Curves $F_n$ intersect at
  $d^{\star}=d_{s}$ which is indicated by a dashed vertical line. Curves $Q_n$ intersect at $\Delta t = \tau_{\kappa}\ln\Big(\bar g / g^{\star}\Big)$. \label{fig:FQ-maps}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{Pn-map}
  \caption{Map $\Pi_{n}:d^{\star}\mapsto d^{\star}$. $\bm{\mathrm{(A)}}$ $\Pi_{n}$ for $n=1,2,3,4$
  at $\gbar=0.6$ $\si{mS/cm^{2}}$. $\bm{\mathrm{(B)}}$ $\Pi_{2}$ with $n=2$ for
  varying $\gbar \approx 0.01, 0.034, 0.3$ $\si{mS/cm^{2}}$. The identity function is
  illustrated by a diagonal line.~\label{fig:Pn-map}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{folds}
  \caption{$\bm{\mathrm{(A)}}$ Fold bifurcation diagrams of stable (continuous curves)
    and unstable (dotted curves) fixed points of $\Pi_{n}$ for varying $n$.
    $\bm{\mathrm{(B)}}$ Cycle periods computed from stable fixed points of $\Pi_{n}$ (blue), and the corresponding periods from stable $n:n$ solutions acquired via numerical integration of the system of ODEs
    (orange).~\label{fig:folds}}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.8\textwidth]{final-bif}
  \caption{Bifurcation diagrams of stable $n:n$ solutions computed analytically from
    fixed points of $\Pi_n$ and plotted on the respective intervals of $\gbar\in
    \big[\gbar_{\mathcal{L}}(n),\gbar_{\mathcal{R}}(n)\big]$ (blue), and computed from numerical integrations of the ODEs (orange).~\label{fig:final-bif}}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: